Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid

### 2016-02-05 Appendix.tex
appendix+V-VI

\bibliographystyle{IEEEtran}  
\bibliography{Article}

\end{document}

>>>>

\subsection{Utility function}
\label{appendix_utility}

We have the following utility function (see main article) that attributes a real value to any coalition S depending on the best contract value that S can propose given the grid policy $ (P^{MIN},\phi)$ :

\begin{equation}
	U(S) = \frac{1}{|S|^{\alpha}} \frac{ P_S^{CRCT \star} }{P^{MAX}}
\end{equation}

$ U $ depends on the specific agents within the coalition S because $ P_S^{CRCT \star} $ is directly impacted by the correlations between the agents. $ U $ is also sensitive to the size of the coalition S through the term $ \frac{1}{|S|^{ \alpha } } $. The purpose of this term is to introduce in the utility function the following tradeoff : as a coalition grows, its contract value $ P_S^{CRCT \star} $ tends to grow as well but the cost of coordinating all the agents also becomes larger. The parameter $ \alpha $ is used in order to control to what extend the size of the coalition impacts its utility.

Knowing what value to use for $ \alpha $ is not trivial and is important since it influences the shape of the utility function. For some range of values, the utility might not even be a concave function of the coalition size. Given a situation with N agents and a desired number of coalitions $ N_{COAL}$, the algorithm \ref{alg:algo1} will start with $ N_{COAL} $ seeds and increase their sizes as long as the utilities are improved. The order through which coalitions incorporate new prosumers is randomized but leads generally to relatively balanced coalitions. We denote by $ \bar{N} $ the desired approximate final size of the coalitions, i.e we target a utility function that is concave in the coalition size with a maximum at $ \bar{N} $ (in the article, $ \bar{N} $ is fixed at $\frac{N}{N_{COAL}}$). We propose to obtain such utility function through tunning $\alpha $. Ideally, we seek to express $ \alpha $ as a function of $ \bar{N} $ and the agents power distributions and correlations.
\\
In the general case, neither the power distributions nor the correlation structure have special forms which makes obtaining an analytical expression for $ \alpha $ difficult. Therefore we use two simplifying assumptions in order to have an estimate for $ \alpha $. First, we approximate the power distributions by normal distributions, such that the optimal contract value can be written in the following way :

\begin{equation}
	P_S^{CRCT \star } = \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi)
\end{equation}

With :

\begin{equation}
	\left\{ \begin{array}{lll} \mu_{S} = \sum_{i \in S} \mu_i \\ \sigma_{S} = \sqrt{ \sum_{i \in S} \sigma_i^2 + 2 \sum_{i} \sum_{j} \rho_{ij} \sigma_i \sigma_j } \end{array} \right.
\end{equation}

Therefore, we can write the utility function as :

\begin{equation}
	U(S) = \frac{1}{|S|^{\alpha} P^{MAX} } \left[ \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \right]
\end{equation}

As $ \mu_S $ and $ \sigma_S $ depends on the agents in S (and consequently on the size of S), we make a second simplifying assumption :

\begin{equation}
	\left\{ \begin{array}{lll} \forall i \in \mathcal{A}\ \mu_i = \bar{ \mu } = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i  \\ \forall i \in \mathcal{A}\ \sigma_i = \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i \\ \forall i,j \in \mathcal{A}^2\  \rho_{ij} = \bar{ \rho } = \frac{2}{N(N-1)} \sum_{i \in \mathcal{A}} \sum_{j \in \mathcal{A}} \rho_{ij} \end{array} \right.
\end{equation}

That is, any agent is replaced by a mean approximation. In these conditions, we have :

\begin{equation}
	\left\{ \begin{array}{lll} 
				\mu_S = |S| \bar{\mu} \\
				\sigma_S = \bar{\sigma} \sqrt{|S| \left[ 1 + \bar{\rho} (|S|-1) \right] }	
	\end{array} \right.
\end{equation}

And the utility function becomes :

\begin{equation}
	U(S) = \frac{ \bar{ \mu } }{ |S|^{\alpha -1} P^{MAX} } - \frac{ \sqrt(2) \bar{ \sigma } }{ |S|^{\alpha} P^{MAX} } \sqrt{|S| + \bar{ \rho } |S| ( |S| - 1 ) }
\end{equation}

$ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $. Depending on $ \alpha $, the utility function can have very different shapes.
\\
We select $\alpha^{\star} $ such that :

\begin{equation}
	\left\{ \begin{array}{lll} \left| \frac{ \partial U }{ \partial |S| } \right|_{|S|=\bar{N}} = 0 \\ \left| \frac{ \partial^2 U }{ \partial |S|^2 } \right|_{\alpha = \alpha^{\star}} \leq 0 \end{array} \right.
\end{equation}

Solving these equations yields the following expression :

\begin{equation}
	\begin{tiny}
		\alpha^{\star} ( \bar{N} ) = \dfrac{0.7 \bar{\sigma}(\bar{\rho}-1)erf^{-1}(2 \phi - 1)}{\bar{\mu}\sqrt{\bar{N}(\bar{\rho}\bar{N}-\bar{\rho}+1)}+1.4 \bar{\sigma} erf^{-1} (2 \phi -1 ) (\bar{\rho}\bar{N}-\bar{\rho} + 1)} 
	\end{tiny}
\end{equation}

Since $ \alpha^{\star} ( \bar{N} ) $ is obtained for a simplified case derived from the real situation, it should be considered as a soft constraint on the coalitions sizes.

Using such an approach on 100 prosumers examples with $N_{COAL} = 4 $ and $\bar{N}=25$ results in coalitions which sizes are comprised between 17 and 34 agents.

\bibliographystyle{IEEEtran}  
\bibliography{Article}
\end{document}

### 2016-02-05 Article.tex
appendix+V-VI

In the general case where distributions and correlation structures have no special form, finding an analytical expression for $ \alpha $ appears complicated. To overcome this problem, we express $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $. In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix for more details). We select $ \alpha^{\star} $ such that :

@@

Which yields to the following expression for $ \alpha^{\star} $ (see Appendix) :

@@

Figure \ref{fig:mean_approx} shows how $ \alpha^{\star} $ and the utility function evolves according to the mean size of the coalitions $ \bar{N} $. As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs. $ \mathcal{U} $ is rather designed as a quality measure for the coalitions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section VI: Coalition Formation
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coalition Formation}
\label{sec:forming}
Section \ref{sec:notations} explained how contract values for the coalitions are computed, and in section \ref{sec:utility} we related this quantity to the utility and gains of a coalition. Since the computation time of this quantity is not negligible, we derive in this section the heuristic we used to form the coalitions structure.

>>>>

In the general case where distributions and correlation structures have no special form, finding an analytical expression for $ \alpha $ appears complicated. To overcome this problem, we seek an approximation for $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $. In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix for more details). We then select $ \alpha^{\star} $ such that :

@@

Which leads to the following expression for $ \alpha^{\star} $ (see Appendix) :

@@

Figure \ref{fig:mean_approx} shows how $ \alpha^{\star} $ and the utility function evolves according to the mean size of the coalitions $ \bar{N} $. Given a policy $ (\phi, P^{MIN} ) $ and a pool of N prosumers, we are now able to quantify the quality of any coalition S.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section VI: Coalition Formation
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coalition Formation}
\label{sec:forming}
We aim at forming $ N_{COAL} $ coalitions $ \{ S_1;S_2;...;S_{N_{COAL}} \} $ such that the global utility $ U = \sum_{i = 1}^{N_{COAL}} \mathcal{U}(S_i) $ is maximized. Since achieving such a goal is a well-known hard problem due to the combinatorial explosion in the number of possible aggregations, we propose in this section a greedy heuristic using decorrelation graphs to solve the problem.
